\hfill\break \hfill\break \hfill\break \vspace{1cm}Customer dropout membership\footnote{Corresponding address: \href{mailto:sobreiro@esdrm.ipsantarem.pt}{\nolinkurl{sobreiro@esdrm.ipsantarem.pt}}. Quality of Life Research Centre, Polytechnic Institute of Santarém, Portugal}\vspace{0.5cm}

### 2021-10-24 article/articleCustomerDropoutMembership.tex
upd

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}
\title{\hfill\break
\hfill\break
\hfill\break
\vspace{1cm}Customer dropout membership\footnote{Corresponding address: \href{mailto:sobreiro@esdrm.ipsantarem.pt}{\nolinkurl{sobreiro@esdrm.ipsantarem.pt}}. The current template adapts part of the Rmd code by \href{https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown}{Paul C. Bauer}, Mannheim Centre for European Social Research.}\vspace{0.5cm}}
\author{Pedro Sobreiro, Javier Berrocal, Domingos Martinho, José Garcia Alonso\\
~\\
Extremadura University\\}
\date{\hfill\break
\hfill\break
25 julho, 2021\\
~\\}

@@

\begin{itemize}
\tightlist
\item
\end{itemize}

@@

\clearpage

\renewcommand{\baselinestretch}{0.5}\normalsize

\renewcommand{\baselinestretch}{1.1}\normalsize

\clearpage

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(dplyr)}
\FunctionTok{library}\NormalTok{(dlookr)}
\FunctionTok{library}\NormalTok{(ggplot2)}

\CommentTok{\#eda\_report(nlswork,output\_dir  = }
\CommentTok{\#   "C:/Users/mangelo.EEG/Documents/GitHub/prjs/reports/",}
\CommentTok{\#   output\_file = "eda\_report.pdf")}

\DocumentationTok{\#\# The data}

\FunctionTok{names}\NormalTok{(airquality)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "Ozone"   "Solar.R" "Wind"    "Temp"    "Month"   "Day"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#summary(nlswork)}

\DocumentationTok{\#\# Missing values}
\FunctionTok{library}\NormalTok{(visdat)}
\FunctionTok{vis\_dat}\NormalTok{(airquality)}
\end{Highlighting}
\end{Shaded}

\includegraphics{articleCustomerDropoutMembership_files/figure-latex/Stats1-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(naniar)}
\FunctionTok{vis\_miss}\NormalTok{(airquality)}
\end{Highlighting}
\end{Shaded}

\includegraphics{articleCustomerDropoutMembership_files/figure-latex/Stats1-2.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{gg\_miss\_upset}\NormalTok{(airquality)}
\end{Highlighting}
\end{Shaded}

\includegraphics{articleCustomerDropoutMembership_files/figure-latex/Stats1-3.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# GRAPHS}
\NormalTok{dplyr}\SpecialCharTok{::}\FunctionTok{glimpse}\NormalTok{(cars}\SpecialCharTok{$}\NormalTok{Ozone)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  NULL
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{d }\OtherTok{\textless{}{-}} \FunctionTok{density}\NormalTok{(airquality}\SpecialCharTok{$}\NormalTok{Temp)}
\FunctionTok{plot}\NormalTok{(d)}
\end{Highlighting}
\end{Shaded}

\includegraphics{articleCustomerDropoutMembership_files/figure-latex/Stats1-4.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot}\NormalTok{(airquality}\SpecialCharTok{$}\NormalTok{Ozone, airquality}\SpecialCharTok{$}\NormalTok{Solar.R)}
\end{Highlighting}
\end{Shaded}

\includegraphics{articleCustomerDropoutMembership_files/figure-latex/Stats1-5.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(airquality, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Ozone, }\AttributeTok{y =}\NormalTok{ Solar.R)) }\SpecialCharTok{+}
\FunctionTok{geom\_miss\_point}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{articleCustomerDropoutMembership_files/figure-latex/Stats1-6.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ggplot}\NormalTok{(airquality, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Wind, }\AttributeTok{y =}\NormalTok{ Temp)) }\SpecialCharTok{+}
\FunctionTok{geom\_miss\_point}\NormalTok{() }\SpecialCharTok{+}
\FunctionTok{facet\_wrap}\NormalTok{(}\FunctionTok{vars}\NormalTok{(Month))}
\end{Highlighting}
\end{Shaded}

\includegraphics{articleCustomerDropoutMembership_files/figure-latex/Stats1-7.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{stats }\OtherTok{\textless{}{-}} \FunctionTok{summary}\NormalTok{(airquality}\SpecialCharTok{$}\NormalTok{Temp)}
\NormalTok{stats}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##   56.00   72.00   79.00   77.88   85.00   97.00
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{describe}\NormalTok{(airquality)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 6 x 26
##   variable     n    na   mean    sd se_mean   IQR skewness kurtosis   p00   p01
##   <chr>    <int> <int>  <dbl> <dbl>   <dbl> <dbl>    <dbl>    <dbl> <dbl> <dbl>
## 1 Ozone      116    37  42.1  33.0    3.06   45.2  1.24       1.29    1    4.3 
## 2 Solar.R    146     7 186.   90.1    7.45  143   -0.428     -0.968   7   10.2 
## 3 Wind       153     0   9.96  3.52   0.285   4.1  0.348      0.111   1.7  2.56
## 4 Temp       153     0  77.9   9.47   0.765  13   -0.378     -0.404  56   57   
## 5 Month      153     0   6.99  1.42   0.115   2   -0.00239   -1.30    5    5   
## 6 Day        153     0  15.8   8.86   0.717  15    0.00265   -1.20    1    1   
## # ... with 15 more variables: p05 <dbl>, p10 <dbl>, p20 <dbl>, p25 <dbl>,
## #   p30 <dbl>, p40 <dbl>, p50 <dbl>, p60 <dbl>, p70 <dbl>, p75 <dbl>,
## #   p80 <dbl>, p90 <dbl>, p95 <dbl>, p99 <dbl>, p100 <dbl>
\end{verbatim}

\hypertarget{experimental-results}{%
\section{Experimental Results}\label{experimental-results}}

\hypertarget{data-description}{%
\subsection{Data description}\label{data-description}}

\begin{verbatim}
##  [1] "Sócio"               "dataAdesao"          "ano"                
##  [4] "dataNascimento"      "idade"               "sexo"               
##  [7] "estadoCivil"         "categoria"           "quotaMensal"        
## [10] "profissao"           "codPostal"           "ultimaQuota"        
## [13] "ultimoPagamento"     "valorTotal"          "totalJogos"         
## [16] "jogosEpoca"          "diasUltimoPagamento" "mesesUP"            
## [19] "abandonou"           "anosSocio"           "idaEstadio"         
## [22] "escaloesTotalJogos"  "mes"
\end{verbatim}

\begin{verbatim}
##  [1] "num_socio"                 "dt_inscription"           
##  [3] "year"                      "birth_date"               
##  [5] "age"                       "sex"                      
##  [7] "marital_status"            "category"                 
##  [9] "monthly_fee"               "occupation"               
## [11] "zip_code"                  "dt_last_invoice"          
## [13] "dt_last_payment"           "total_amount"             
## [15] "total_matches"             "season_matches"           
## [17] "days_since_last_payment"   "months_since_last_payment"
## [19] "dropout"                   "years_membership"         
## [21] "stadium_access"            "quart_stadium_entries"    
## [23] "inscription_month"
\end{verbatim}

\begin{verbatim}
## tibble [25,316 x 14] (S3: tbl_df/tbl/data.frame)
##  $ year                     : num [1:25316] 1944 1944 1945 1945 1945 ...
##  $ age                      : num [1:25316] 83 88 73 97 97 91 88 95 88 78 ...
##  $ sex                      : chr [1:25316] "M" "M" "M" "M" ...
##  $ marital_status           : chr [1:25316] "casado" "solteiro" "nao definido" "casado" ...
##  $ monthly_fee              : num [1:25316] 10 10 10 5 10 5 5 5 10 10 ...
##  $ total_amount             : num [1:25316] 1906 1906 1553 790 1466 ...
##  $ total_matches            : num [1:25316] 0 0 0 0 0 20 74 0 154 0 ...
##  $ season_matches           : num [1:25316] 0 0 0 0 0 0 0 0 6 0 ...
##  $ months_since_last_payment: num [1:25316] 3 3 36 8 35 4 41 40 4 2 ...
##  $ dropout                  : num [1:25316] 0 0 1 0 1 0 1 1 0 0 ...
##  $ years_membership         : num [1:25316] 74 74 73 73 73 73 73 73 73 72 ...
##  $ stadium_access           : num [1:25316] 0 0 0 0 0 1 1 0 1 0 ...
##  $ quart_stadium_entries    : chr [1:25316] "ate 1" "ate 1" "ate 1" "ate 1" ...
##  $ inscription_month        : num [1:25316] 10 10 8 9 9 12 1 1 2 4 ...
\end{verbatim}

\includegraphics{articleCustomerDropoutMembership_files/figure-latex/get_data-1.pdf}

Table \ref{tab:tab1cars} shows data's summary statistics.\footnote{You can reference the table as \ref{tab1cars}.} \texttt{stargazer()} is
and excellent solution to export outputs.

Teste \ref{tab:summarytable}

@@

\caption{\label{tab:summarytable}Summary statistics}
\centering
\begin{tabular}[t]{ll}
\toprule
Characteristic & N = 25,316\\
\midrule
year & 2,007 (11)\\
age & 27 (20)\\
sex & \\
\hspace{1em}F & 32\%\\
\hspace{1em}M & 68\%\\
\addlinespace
marital\_status & \\
\hspace{1em}casado & 20\%\\
\hspace{1em}nao definido & 30\%\\
\hspace{1em}outro & 2.0\%\\
\hspace{1em}solteiro & 48\%\\
\addlinespace
monthly\_fee & \\
\hspace{1em}0 & <0.1\%\\
\hspace{1em}1 & 32\%\\
\hspace{1em}2.5 & 28\%\\
\hspace{1em}5 & 3.4\%\\
\addlinespace
\hspace{1em}6 & 12\%\\
\hspace{1em}10 & 24\%\\
total\_amount & 316 (494)\\
total\_matches & 27 (46)\\
season\_matches & 2.2 (4.1)\\
\addlinespace
months\_since\_last\_payment & 19 (32)\\
dropout & 22\%\\
years\_membership & 11 (11)\\
stadium\_access & 40\%\\
quart\_stadium\_entries & \\
\addlinespace
\hspace{1em}1 a 21 & 10\%\\
\hspace{1em}21 a 56 & 9.8\%\\
\hspace{1em}56 a 105 & 10.0\%\\
\hspace{1em}ate 1 & 60\%\\
\hspace{1em}mais 105 & 10.0\%\\
\addlinespace
inscription\_month & 6.9 (3.4)\\
\bottomrule
\multicolumn{2}{l}{\rule{0pt}{1em}\textsuperscript{1} Mean (SD); \%}\\
\end{tabular}
\end{table}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#from pysurvival.utils.display import correlation\_matrix}
\ImportTok{import}\NormalTok{ pandas }\ImportTok{as}\NormalTok{ pd}
\ImportTok{import}\NormalTok{ numpy }\ImportTok{as}\NormalTok{ np}

\NormalTok{col }\OperatorTok{=}\NormalTok{ [}\StringTok{\textquotesingle{}sex\textquotesingle{}}\NormalTok{,}\StringTok{\textquotesingle{}marital\_status\textquotesingle{}}\NormalTok{,}\StringTok{\textquotesingle{}quart\_stadium\_entries\textquotesingle{}}\NormalTok{]}

\NormalTok{df\_members }\OperatorTok{=}\NormalTok{ r.df\_members }\CommentTok{\#copy r dataframe to python}

\NormalTok{df\_members }\OperatorTok{=}\NormalTok{ pd.get\_dummies(df\_members, columns}\OperatorTok{=}\NormalTok{col,drop\_first}\OperatorTok{=}\VariableTok{True}\NormalTok{)}

\CommentTok{\# Creating the time and event columns}
\NormalTok{time\_column }\OperatorTok{=} \StringTok{\textquotesingle{}years\_membership\textquotesingle{}}
\NormalTok{event\_column }\OperatorTok{=} \StringTok{\textquotesingle{}dropout\textquotesingle{}}

\CommentTok{\# Extracting the features}
\NormalTok{features }\OperatorTok{=}\NormalTok{ np.setdiff1d(df\_members.columns, [time\_column, event\_column] ).tolist()}

\CommentTok{\#correlation\_matrix(df\_members, figure\_size=(10,10), text\_fontsize=6)}
\end{Highlighting}
\end{Shaded}

The average age in our data is 27.3.

\hypertarget{sec:tables}{%
\section{Tables}\label{sec:tables}}

R Markdown PDF is now able to produce good tables with our output. For
\texttt{stargazer} the label is contained in the function, while for \texttt{kable} it's
contained in the chunk name.

\hypertarget{stargazer-summary-and-regression-tables}{%
\subsection{stargazer(): Summary and regression tables}\label{stargazer-summary-and-regression-tables}}

Table \ref{tab2} reports regression outputs. Name the models as you can refer
to their names in the text (M1, M2, M3).

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(stargazer)}
\NormalTok{model1 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(speed }\SpecialCharTok{\textasciitilde{}}\NormalTok{ dist, }\AttributeTok{data =}\NormalTok{ cars)}
\NormalTok{model2 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(speed }\SpecialCharTok{\textasciitilde{}}\NormalTok{ dist, }\AttributeTok{data =}\NormalTok{ cars)}
\NormalTok{model3 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(dist }\SpecialCharTok{\textasciitilde{}}\NormalTok{ speed, }\AttributeTok{data =}\NormalTok{ cars)}
\FunctionTok{stargazer}\NormalTok{(model1, model2, model3,}
          \AttributeTok{title =} \StringTok{"Regression table with stargazer"}\NormalTok{,}
          \AttributeTok{label =} \StringTok{"tab2"}\NormalTok{,}
          \AttributeTok{table.placement =} \StringTok{"h"}\NormalTok{,}
          \AttributeTok{column.labels =} \FunctionTok{c}\NormalTok{(}\StringTok{"M1"}\NormalTok{, }\StringTok{"M2"}\NormalTok{, }\StringTok{"M3"}\NormalTok{),}
          \AttributeTok{model.numbers =} \ConstantTok{FALSE}\NormalTok{,}
          \AttributeTok{header =} \ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[h] \centering 
  \caption{Regression table with stargazer} 
  \label{tab2} 
\begin{tabular}{@{\extracolsep{5pt}}lccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:}} \\ 
\cline{2-4} 
\\[-1.8ex] & \multicolumn{2}{c}{speed} & dist \\ 
 & M1 & M2 & M3 \\ 
\hline \\[-1.8ex] 
 dist & 0.166$^{***}$ & 0.166$^{***}$ &  \\ 
  & (0.017) & (0.017) &  \\ 
  & & & \\ 
 speed &  &  & 3.932$^{***}$ \\ 
  &  &  & (0.416) \\ 
  & & & \\ 
 Constant & 8.284$^{***}$ & 8.284$^{***}$ & $-$17.579$^{**}$ \\ 
  & (0.874) & (0.874) & (6.758) \\ 
  & & & \\ 
\hline \\[-1.8ex] 
Observations & 50 & 50 & 50 \\ 
R$^{2}$ & 0.651 & 0.651 & 0.651 \\ 
Adjusted R$^{2}$ & 0.644 & 0.644 & 0.644 \\ 
Residual Std. Error (df = 48) & 3.156 & 3.156 & 15.380 \\ 
F Statistic (df = 1; 48) & 89.567$^{***}$ & 89.567$^{***}$ & 89.567$^{***}$ \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\hypertarget{figures}{%
\section{Figures}\label{figures}}

\hypertarget{graphs-with-r}{%
\subsection{Graphs with R}\label{graphs-with-r}}

You can insert figures like this. One would like to produce and insert them on
the fly in the \texttt{.rmd} file. Figure \ref{fig:fig-1} is such an example.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot}\NormalTok{(cars}\SpecialCharTok{$}\NormalTok{speed, cars}\SpecialCharTok{$}\NormalTok{dist)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[h]

{\centering \includegraphics{articleCustomerDropoutMembership_files/figure-latex/Figures 1, fig-1-1}

}

\caption{Scatterplot of Speed and Distance}(\#fig:Figures 1, fig-1)
\end{figure}

However, in some cases it does not work.

\hypertarget{example-ggplot2-graphs}{%
\subsection{Example: ggplot2 graphs}\label{example-ggplot2-graphs}}

See the \texttt{ggplot2} output reported in Figure \ref{fig:fig-2}.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}
\title{\hfill\break
\hfill\break
\hfill\break
\vspace{1cm}Customer dropout membership\footnote{Corresponding address: \href{mailto:sobreiro@esdrm.ipsantarem.pt}{\nolinkurl{sobreiro@esdrm.ipsantarem.pt}}. Quality of Life Research Centre, Polytechnic Institute of Santarém, Portugal}\vspace{0.5cm}}
\author{Pedro Sobreiro, Javier Berrocal, Domingos Martinho, José Garcia Alonso\\
\strut \\
Extremadura University\\}
\date{\hfill\break
\hfill\break
9 outubro, 2021\\
\strut \\}

@@

@@

Churn dropout prediction is a problem being addressed supported in the idea that
the customers database is the most valuable asset that the organizations possess
(\protect\hyperlink{ref-Athanassopoulos_2000}{Athanassopoulos 2000}), which requires determining customers that will attrite
(\protect\hyperlink{ref-alboukaey_dynamic_2020}{Alboukaey, Joukhadar, and Ghneim 2020a}). Dropout implies in contractual business that the
customer needs to renew their contracts to continue its usage
(\protect\hyperlink{ref-Ascarza_Hardie_2013}{Ascarza and Hardie 2013}).

However, in contractual settings the customer dropout represents an explicit
ending of a relationship more punitive than non contractual settings
(\protect\hyperlink{ref-risselada_staying_2010}{Risselada, Verhoef, and Bijmolt 2010}) This has implications to the profitability of the
organizations increasing marketing costs and reducing sales
(\protect\hyperlink{ref-amin_customer_2017}{Amin et al. 2017}).

The anticipation of the dropout allows the development of countermeasures to
reduce customer churn. Several studies address the problem related to customer
retention trying to improve the profitability (\protect\hyperlink{ref-coussement_improving_2009}{Coussement and Van den Poel 2009}; \protect\hyperlink{ref-devriendt_why_2019}{Devriendt, Berrevoets, and Verbeke 2019}; \protect\hyperlink{ref-garcia_intelligent_2017}{García, Nebot, and Vellido 2017})

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{If an organization can predict a possible dropout and}
\NormalTok{develop countermeasures to avoid desertion, they can avoid customer defections that }
\NormalTok{lead to a loss of money. Reichheld (1996) evidenced that reducing dropout rates by 5\% }
\NormalTok{(e.g., from 15\% to 10\% per year) could represent an increase in prots up to double, as }
\NormalTok{acquiring new customers costs 5 to 6 times more than retaining existing ones (Bhattacharya, }
\NormalTok{1998). Existing organizations are addressing this problem by shifting their target from }
\NormalTok{capturing new customers to preserving existing ones (García et al., 2017),}
\NormalTok{as investments in retention strategies have higher returns than acquisitions (Coussement }
\NormalTok{\& Van den Poel, 2009). The importance of customer retention to maintain organizational protability (Devriendt}
\NormalTok{et al., 2019) leads to the problem of how to quantify the nancial impact of customer retention actions}
\NormalTok{under the assumption that the organization goal should be related to the increase the lifetime of the}
\NormalTok{customer to increase their prots. The customer lifetime value (CLV) allows us to measure this, as itr}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Address the global problem of customer dropout
\item
  The identification of approaches to predict dropout requires more than only,
  address the prediction accuracy such as \ldots{} place existing studies
  addressing this\ldots{}
\item
  A lot of effort has been placed testing the accuracy of existing algorithms,
  in this study we try to fill this gap and explore also a balancing between
  the model interpretability, accuracy, and the investment required.
\item
  Dropout prediction problem related to the timings
\end{enumerate}

The approaches normally employing use a dependent variable representing dropout
or non-dropout, without considering a dynamic perspective that the dropout risk
changes over time (\protect\hyperlink{ref-Alboukaey_dynamic_2020}{Alboukaey, Joukhadar, and Ghneim 2020b}). The survival models try to solve
this limitation (\protect\hyperlink{ref-routh_estimating_2020}{Routh, Roy, and Meyer 2020}) capturing a temporal dimension of the
customer dropout (\protect\hyperlink{ref-perianez_churn_2016}{Perianez et al. 2016}). \protect\hyperlink{ref-perianez_churn_2016}{Perianez et al.} (\protect\hyperlink{ref-perianez_churn_2016}{2016}) used survival
analysis to predict also when the dropout will occur.

Other studies proposed also the integration of several algorithms to improve the
performance in the prediction of the dropout such the usage of clusters combined
with churn prediction (\protect\hyperlink{ref-gok_case_2015}{Gök, Özyer, and Jida 2015}; \protect\hyperlink{ref-hung_applying_2006}{Hung, Yen, and Wang 2006}; \protect\hyperlink{ref-vijaya_sivasankar_2019}{Vijaya and Sivasankar 2019}). The approach relies in the assumption that combining
the customers in different clusters allows the improvement of the prediction
accuracy. \protect\hyperlink{ref-vijaya_sivasankar_2019}{Vijaya and Sivasankar} (\protect\hyperlink{ref-vijaya_sivasankar_2019}{2019}) suggested the adoption hybrid models combining
more than one classier are achieving increased performance compared to those
using single classifiers.

There are several challenges around the timing related to dropout, or
considering the dynamic behavior of the customer in the intent to drop out
\[@alboukaey_dynamic_2020\]. The importance of understanding when dropout will
occur and the risk when discarding the temporal perspective of the problem seems
to be an element that should be addressed. Few studies considered this
\[@perianez_churn_2016; @burez_separating_2008\]. This shows an opportunity to
address the importance of the timeframe and its influence on the efficiency of
the model and also evalute if the combination of clusters could improve the
performance.

In this study, we adopt random survival forests which have never been used in
understanding factors affecting membership in a sport club using existing data
in a Sport Club. The analysis is based on the use of random survival forests in
the presence of covariates that do not necessarily satisfy the PH assumption.
Additionly we also propose a new approach combining clusters with survival
analysis.

??? Add interpretability layer

Random Survival Forests does not make the proportional hazards assumption
(\protect\hyperlink{ref-Ehrlinger_2016}{Ehrlinger 2016}) and has the flexibility to model survivor curves that are of
dissimilar shapes for contrasting groups of subjects. Random Survival Forest is
an extension of Random Forest allowing efficient non-parametric analysis of time
to event data (\protect\hyperlink{ref-Breiman_2001}{Breiman 2001}). This characteristics allow us to surpass the Cox
Regression limitation of the proportional hazard assumption, requiring to
exclude variables which not fulfill the model assumption. It was shown by
\protect\hyperlink{ref-Breiman_2001}{Breiman} (\protect\hyperlink{ref-Breiman_2001}{2001}) that ensemble learning can be further improved by injecting
randomization into the base learning process - a method called Random Forests.

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

Dropout is a binary value where one represent churn and zero not churn. The
dropout happens when a member does not have a payment \ldots{}

The model performance was determined with the concordance probability (C-index),
Brier Score (BS) and Mean Absolute Error (MAE) (\protect\hyperlink{ref-wangmachine2017}{Wang, Li, and Reddy 2017}). The feature
importance was determined calculating the difference between the true class
label and noised data (\protect\hyperlink{ref-Breiman_2001}{Breiman 2001}).

\hypertarget{dataset}{%
\subsection{Dataset}\label{dataset}}

Table \ref{tab:summarytable} shows data's summary statistics. The average age
is 27.3 ±
20.1, the members have an attendance of
27 ±
45.8 with a membership of
11 ±
10.9 years.

@@

\caption{\label{tab:summarytable}Summary statistics of features used}
\centering
\begin{tabular}[t]{ll}
\toprule
Characteristic & N = 25,316\\
\midrule
Age in years, Mean (SD) & 27 (20)\\
Male or female, \% & \\
\hspace{1em}F & 32\%\\
\hspace{1em}M & 68\%\\
Single, married and other., \% & \\
\addlinespace
\hspace{1em}casado & 20\%\\
\hspace{1em}nao definido & 30\%\\
\hspace{1em}outro & 2.0\%\\
\hspace{1em}solteiro & 48\%\\
monthly\_fee, \% & \\
\addlinespace
\hspace{1em}0 & <0.1\%\\
\hspace{1em}1 & 32\%\\
\hspace{1em}2.5 & 28\%\\
\hspace{1em}5 & 3.4\%\\
\hspace{1em}6 & 12\%\\
\addlinespace
\hspace{1em}10 & 24\%\\
total\_amount, Mean (SD) & 316 (494)\\
total\_matches, Mean (SD) & 27 (46)\\
season\_matches, Mean (SD) & 2.2 (4.1)\\
months\_since\_last\_payment, Mean (SD) & 19 (32)\\
\addlinespace
dropout, \% & 22\%\\
years\_membership, Mean (SD) & 11 (11)\\
stadium\_access, \% & 40\%\\
quart\_stadium\_entries, \% & \\
\hspace{1em}1 a 21 & 10\%\\
\addlinespace
\hspace{1em}21 a 56 & 9.8\%\\
\hspace{1em}56 a 105 & 10.0\%\\
\hspace{1em}ate 1 & 60\%\\
\hspace{1em}mais 105 & 10.0\%\\
inscription\_month, Mean (SD) & 6.9 (3.4)\\
\bottomrule
\end{tabular}
\end{table}

Figure \ref{fig:membershipyear} shows the distribution of the dropout
considering the number of years of membership.

\begin{figure}
\centering
\includegraphics{articleCustomerDropoutMembership_files/figure-latex/membershipyear-1.pdf}
\caption{\label{fig:membershipyear}Number of members by year}
\end{figure}

\hypertarget{model-construction}{%
\subsection{Model construction}\label{model-construction}}

Address the model construction\ldots{} the categorical variables \(sex\),
\(marital\_status\) and \(quart\_stadium\_entries\) where converted to dummy
variables.

The random survival forest was developed using the package PySurvival
(\protect\hyperlink{ref-Fotso_others_2019}{Fotso et al. 2019}). The most relevant variables predicting the dropout are
analysed using the log-rank test. The metric variables are transformed to
categorical using the quartiles to provide a statistical comparison of groups.
The survival analysis was conducted using the package Lifelines
(\protect\hyperlink{ref-Davidson-Pilon_2021}{Davidson-Pilon 2021}).

PySurvival is an open source python package for Survival Analysis modeling - the
modeling concept used to analyze or predict when an event is likely to happen.
It is built upon the most commonly used machine learning packages such NumPy,
SciPy and PyTorch. PySurvival is compatible with Python 2.7-3.7

\hypertarget{survival-trees-based-model}{%
\subsubsection{Survival trees based model}\label{survival-trees-based-model}}

In this model\ldots{} The survival trees based model uses pysurvival random forest

Removed the variables with greater correlations \(total\_matches\) and
\(quart\_stadium\_entries\)

\includegraphics{articleCustomerDropoutMembership_files/figure-latex/removeVar-1.pdf}

\begin{verbatim}
##         age  ...  quart_stadium_entries_mais 105
## 0      83.0  ...                               0
## 1      88.0  ...                               0
## 2      73.0  ...                               0
## 3      97.0  ...                               0
## 4      97.0  ...                               0
## ...     ...  ...                             ...
## 25311   7.0  ...                               0
## 25312   8.0  ...                               0
## 25313   2.0  ...                               0
## 25314  14.0  ...                               0
## 25315  28.0  ...                               0
## 
## [25316 rows x 14 columns]
\end{verbatim}

\begin{verbatim}
## RandomSurvivalForestModel
\end{verbatim}

\includegraphics{articleCustomerDropoutMembership_files/figure-latex/createModel-3.pdf} \includegraphics{articleCustomerDropoutMembership_files/figure-latex/createModel-4.pdf}

\begin{verbatim}
## {'root_mean_squared_error': 57.814697481046416, 'median_absolute_error': 18.96621852166039, 'mean_absolute_error': 38.55719494742953}
\end{verbatim}

Table \ref{tab:summarytable2} shows variables importance.

\caption{\label{tab:summarytable2}Summary statistics of features used}
\centering
\begin{tabular}[t]{lrr}
\toprule
feature & importance & pct\_importance\\
\midrule
dropout & 6.0669405 & 0.2528415\\
months\_since\_last\_payment & 4.9581586 & 0.2066327\\
season\_matches & 2.3851555 & 0.0994020\\
total\_amount & 1.8950237 & 0.0789757\\
quart\_stadium\_entries\_21 a 56 & 1.4643279 & 0.0610263\\
\addlinespace
quart\_stadium\_entries\_56 a 105 & 1.3455565 & 0.0560765\\
stadium\_access & 1.2980882 & 0.0540982\\
inscription\_month & 1.2790920 & 0.0533065\\
marital\_status\_nao definido & 1.0339551 & 0.0430904\\
sex\_M & 0.8881565 & 0.0370142\\
\addlinespace
age & 0.7472122 & 0.0311403\\
monthly\_fee & 0.6333690 & 0.0263958\\
marital\_status\_solteiro & -0.2702218 & 0.0000000\\
marital\_status\_outro & -1.0259784 & 0.0000000\\
quart\_stadium\_entries\_mais 105 & -1.2762422 & 0.0000000\\
\addlinespace
years\_membership & -4.1258616 & 0.0000000\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{model-building}{%
\paragraph{Model building}\label{model-building}}

The model was built with with 70\% of the data for training and 30\% for testing.
The survival model parameters where:

The model accuracy is very high in the first years. The prediction is very
similar to the actual value. The absolute error mean of
39 customers.

\hypertarget{survival-trees-based-model-with-clusters}{%
\subsubsection{Survival trees based model with clusters}\label{survival-trees-based-model-with-clusters}}

Here we are will create clusters and developed the optimization within each
cluster\ldots{}

The calculation of he number of clusters used the package mclust (\protect\hyperlink{ref-scrucca2016}{Scrucca et al. 2016})
using the Bayesian Information Criterion (BIC). The model that gives the minimum
BIC score can be selected as the best model (\protect\hyperlink{ref-schwarz1978}{Schwarz 1978}) simplifying the
problem related to choosing the number of components and identifying the
structure of the covariance matrix, based on modelling with multivariate normal
distributions for each component that forms the data set (\protect\hyperlink{ref-akogul2016}{Akogul and Erisoglu 2016}).

In multivariate models are available the following approaches:

\begin{itemize}
\tightlist
\item
  ``EII''spherical, equal volume
\item
  ``EEE''ellipsoidal, equal volume, shape, and orientation
\item
  ``VII''spherical, unequal volume
\item
  ``VVV''ellipsoidal, varying volume, shape, and orientation, which is used as
  default for initialization of EM algorithm
\item
  VVI'': diagonal, varying volume and shape t
\end{itemize}

Estou a ter problemas com o cálculo dos clusters com o BIC\ldots{} estava a aqui a
confirmar os clusters, talvez seja melhor reduzir as variáveis\ldots{} testar a
abordagem aos clusters no artigo dos vinhos\ldots{}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(NbClust)}
\NormalTok{nb }\OtherTok{\textless{}{-}} \FunctionTok{NbClust}\NormalTok{(y, }\AttributeTok{diss=}\ConstantTok{NULL}\NormalTok{, }\AttributeTok{distance =} \StringTok{"euclidean"}\NormalTok{, }
              \AttributeTok{min.nc=}\DecValTok{2}\NormalTok{, }\AttributeTok{max.nc=}\DecValTok{5}\NormalTok{, }\AttributeTok{method =} \StringTok{"kmeans"}\NormalTok{, }
              \AttributeTok{index =} \StringTok{"all"}\NormalTok{, }\AttributeTok{alphaBeale =} \FloatTok{0.1}\NormalTok{)}
\FunctionTok{hist}\NormalTok{(nb}\SpecialCharTok{$}\NormalTok{Best.nc[}\DecValTok{1}\NormalTok{,], }\AttributeTok{breaks =} \FunctionTok{max}\NormalTok{(}\FunctionTok{na.omit}\NormalTok{(nb}\SpecialCharTok{$}\NormalTok{Best.nc[}\DecValTok{1}\NormalTok{,])))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## KMeans(n_clusters=1)
## KMeans(n_clusters=2)
## KMeans(n_clusters=3)
## KMeans(n_clusters=4)
## KMeans(n_clusters=5)
## KMeans(n_clusters=6)
## KMeans(n_clusters=7)
## KMeans()
## KMeans(n_clusters=9)
## KMeans(n_clusters=10)
## KMeans(n_clusters=11)
## KMeans(n_clusters=12)
## KMeans(n_clusters=13)
## KMeans(n_clusters=14)
## KMeans(n_clusters=15)
## KMeans(n_clusters=16)
## KMeans(n_clusters=17)
## KMeans(n_clusters=18)
## KMeans(n_clusters=19)
\end{verbatim}
